---
title: "Class Reference"
author: "Derek Willis"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this a personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

### How to set the working directory

The command to set a working directory is setwd(). For example, this sets the working directory to a directory called data_journalism_2024_fall inside a directory called jour472 located inside your home directory (the ~ is equivalent of /Users/[youruser] on a Mac).

```{r}
setwd("~/jour472/data_journalism_2026_spring")
```

### How to install and load the tidyverse

```{r}
install.packages("tidyverse")
library(tidyverse)
```

```{r}
## the entire group of data (the md churches0 is grouped by their code according to which code has that many churches in it in descending order
md_churches |>
  group_by(NTEE_CD) |>
  summarise(
    count = n()
  ) |>
  arrange(desc(count))
```

```
```{r}
## filter code can shave everything unnecessary off ##
## example below ##
cp_churches <- md_churches |> filter(CITY == "COLLEGE PARK")

## you can also filter out based on a column with a number, like below:
md_churches_indy <- md_churches |> filter(AFFILIATION == 3)
## this would be for the churches' specfic affiliation, the 3 is from the IRS codebook

## the 'select" code doesn't't shave how many rows you see, but how many columns you see!!!!! select doesn't change anything abt the data you're looking at, only which column -- you will probably mainly use the filter command, its much better to filter than select, SELECT is much better when you know specifically what you're looking for ##
```
```{r}
##you can use an ampersand or a | (this would be or) to filter a bit more strategically in the code)

cp_churches_ind <- md_churches |> filter(CITY == "COLLEGE PARK" & AFFILIATION == 3)

nrow(cp_churches_ind)

## now for the OR, which means either one would be true, versus ampersand meaning both need to be true
cp_churches_ind <- md_churches |> filter(CITY == "COLLEGE PARK" | AFFILIATION == 3)

nrow(cp_churches_ind)

##MAKE SURE YOU ARE QUADRUPLE CHECKING WHAT COULD BE WRONG WHEN FILTERING THINGS. humans are imperfect and sometimes things are wrong, meaning soemtiems teh software will be direct according to what was input!!!!!!!!

```
```{r}
##to write dates in code, start with year, then month, then day
```
```{r}
## if you clear out the entire environment (broom icn at top right) and then go to the 'run all' option, you can check to see that its correct from beginning to end

##you can also check individual parts of code, highlight the part you want to see if correct and then go to "run selected"

##most amount is sum##

##ggplot is best way to make plots in r
```

