---
title: "Reverse Engineering Project"
author: "Koree Perry"
date: "3/26/26"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [When Are Half Your Members Going To Be Dead?](https://thescoop.org/religion/reverse-engineering/going-dead.html)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
##Source of data: Ryan Burge - Data Cooperative Election Study

# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

reverse_dead_data <- read_csv("data/cces24_going_dead.csv") |> clean_names()

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [For instance, just 6% of Southern Baptist adults are under the age of 30. That’s no different at all than the Episcopalians. Only about one in five are under the age of 45]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not? -- It took some trial and error to get the results of my analysis. My original answer was 6.7 something, but I decided to round just to make it one number, and it moved it up one percent than the article said. I'm not actually able to confirm because they are two different numbers, but I think I got close to how the data came to be in the article for Southern Baptists. For EPISCOPALIANS: ]

```{r}
# Put code to reverse engineer sentence here
southbap <- reverse_dead_data |>
  filter(trad == "Southern Baptist") |>
  mutate(under30 = age < 30) |>
  arrange(desc(southbap))
 

southbap |>
  summarise(under30_pct = mean(under30) * 100) |> round()
```

```{r}
episco <- reverse_dead_data |>
  filter(trad, "Episcopal") |>
  mutate(under45 = age < 45) |>
  arrange(desc(episco)) 

episco |> 
  summarise(under45_pct = mean(under45) * 100) |> round()
```


# Display results of code below this codeblock

### Sentence 2

* **Sentence text**: [And for 16 of the 20 denominations on this list, Boomers make up at least 40% of their adult membership.]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
## 1946-1964 is baby boomer 
## https://dplyr.tidyverse.org/reference/between.html
# Put code to reverse engineer sentence here
##count is how many total, and total_boom is how many boomers are in each count for each trad

boom <- reverse_dead_data |>
  group_by(trad) |>
  summarise(
    count = n(),
    total_boom = sum(between(birthyr, 1946, 1964))
  )
  
##Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [This should come as a shock to no one who is vaguely aware of American religion — Episcopalians are old. In fact, two-thirds of their adult members have celebrated their 60th birthday. In contrast, just 6% are under the age of thirty.]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
##accidentally pressed the knit button and i don't know if that messed anything up
# Put code to reverse engineer sentence here
episco_old_bday <- reverse_dead_data |>
  filter(str_detect(trad, "Episcopal")) |>
  group_by(age) |>
  summarise(
    pct_at_60 = reverse_dead_data$age(age > 60)
)
# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [ For instance, the average adult member of an Episcopal church is 61 years old. For Southern Baptists, it’s 58. The Assemblies of God are doing a bit better on this metric at 54]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [And for 16 of the 20 denominations on this list, Boomers make up at least 40% of their adult membership.]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-