---
title: "Reverse Engineering Project"
author: "Koree Perry"
date: "3/26/26"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction 

In this notebook, we are reverse engineering the story, [When Are Half Your Members Going To Be Dead?](https://thescoop.org/religion/reverse-engineering/going-dead.html)

## Load libraries

Loading required libraries for this analysis.

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(janitor)
library(lubridate)
```

## Load and Cleaning Data

In this section, describe the source of the data, write a basic data dictionary for data you are working with, and discuss any caveats or issues you discovered working with this data. 

```{r}
##Source of data: Ryan Burge - Data Cooperative Election Study

# Load required data

# Path to data should be loaded from folder "data" i.e. read_csv("data/name_of_data.csv")

reverse_dead_data <- read_csv("data/cces24_going_dead.csv") |> clean_names() |> rename(denom = trad)

# Clean required data and prepare for analysis if needed. 


```

## Sentences to Engineer
In this notebook, we are reverse engineering five sentences from the story.

### Sentence 1

* **Sentence text**: [For instance, just 6% of Southern Baptist adults are under the age of 30. That’s no different at all than the Episcopalians. Only about one in five are under the age of 45]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not? -- It took some trial and error to get the results of my analysis, but I think I got close to how the data came to be in the article for Southern Baptists with 6.74. For EPISCOPALIANS: I don't understand what went wrong when its the exact same code. ]

```{r}
# Put code to reverse engineer sentence here
southbap <- reverse_dead_data |>
  filter(denom == "Southern Baptist") |>
  mutate(under30 = age < 30) |>
  arrange(desc(southbap))
 

southbap |>
  summarise(under30_pct = mean(under30) * 100)
```

```{r}
episco <- reverse_dead_data |>
  filter(denom == "Episcopal") |>
  mutate(under45 = age < 45) |>
  arrange(desc(episco))

episco |> 
  summarise(under45_pct = mean(under45) * 100)
```


# Display results of code below this codeblock

### Sentence 2

* **Sentence text**: [And for 16 of the 20 denominations on this list, Boomers make up at least 40% of their adult membership.]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not? - I was able to figure out how to do it, but the final issue came when I realized I had done the code for the entire list of denominations and not the selected twenty in the article. when I tried to put them in and change/rerun, it wouldn't work, and when I took them out, it wouldn't work again.]

```{r}
## 1946-1964 is baby boomer 
## https://dplyr.tidyverse.org/reference/between.html
# Put code to reverse engineer sentence here
##count is how many total, and total_boom is how many boomers are in each count for each trad

boom <- reverse_dead_data |>
  group_by(denom) |>
  summarise(
    count = n(),
    total_boom = sum(between(birthyr, 1946, 1964), na.rm = TRUE) 
  ) |>
  mutate(
    final_40_pct = (total_boom / count) * 100)

toppercent <- boom |>
  filter(final_40_pct > 40) |>
nrow()
##Display results of code below this codeblock

```

### Sentence 3

* **Sentence text**: [This should come as a shock to no one who is vaguely aware of American religion — Episcopalians are old. In fact, two-thirds of their adult members have celebrated their 60th birthday. In contrast, just 6% are under the age of thirty.]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
##accidentally pressed the knit button and i don't know if that messed anything up
# Put code to reverse engineer sentence here
episco_old_bday <- reverse_dead_data |>
  filter(denom == "Episcopal") |>
  group_by(age) |>
  summarise(
    pct_at_60 = mutate(reverse_dead_data(age > 60))
)
# Display results of code below this codeblock

```

### Sentence 4

* **Sentence text**: [ For instance, the average adult member of an Episcopal church is 61 years old. For Southern Baptists, it’s 58. The Assemblies of God are doing a bit better on this metric at 54]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here
reverse_dead_data |>
  group_by(denom)
  summarise(
    count = n()
  )

# Display results of code below this codeblock

```

### Sentence 5

* **Sentence text**: [And for 16 of the 20 denominations on this list, Boomers make up at least 40% of their adult membership.]
* **Analysis summary**: [Write up two to three sentences describing the results of your analysis.  Were you able to confirm the finding? If not, why not?]

```{r}
# Put code to reverse engineer sentence here

# Display results of code below this codeblock

```

-30-